Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2019-10-29 manuscript/climate_sensitivity_draft.Rmd
changed predicted direction of WD response
following Hoffmann et al. 2011, who found negative response in NC.
WD never had a sig effect, so this has no impact on results/interpretation.

A second fundamental question regarding forest responses to drought is how species' traits -- alone and in interaction with tree size -- influence drought response. Xylem architecture plays a role, with diffuse porous species tending to be more drought sensitive than ring-porous species [@kannenberg_linking_2019] *(Elliot et al. 2015-doi: 10.1111/gcb.13045, Friedrichs et al. 2009-doi:10.1007/s00468-009-0315-2)*, but does not differentiate species beyond broad classes. To link species' level drought responses to fundamental physiological characteristics, it is important to understand how traits shape drought responses. Commonly measured traits including wood density ($WD$) and leaf mass per area ($LMA$) have been linked to drought responses in temperate deciduous forests [@abrams_adaptations_1990; @guerfel_impacts_2009]*(Hoffmann et al. 2011- doi: 10.1111/j.1365-2486.2011.02401.x)* and other forest biomes around the world (*Greenwood et al. 2007- doi: 10.1111/ele.12748*). However, the direction of response is not always consistent; for example, wood density correlated negatively with drought tree drought performance  in a broadleaf deciduous forest in the southeast US *(Hoffmann et al. 2011)* but positively across a global scale (*Greenwood et al. 2007- doi: 10.1111/ele.12748*). Leaf hydraulic traits with more direct linkage to plant hydraulic function, including leaf area shrinkage upon dessication ($PLA_{dry}$; [@scoffoni_leaf_2014]) and turgor loss point ($\pi_{tlp}$)--i.e., the water potential at which leaf wilting occurs [@bartlett_correlations_2016], are emerging as traits with potential to explain greater variation in plant distribution and function than the more commonly-measured traits such as $WD$ and $LMA$ [@medeiros_extensive_2019]. The ability of these hydraulic traits to tree performance under drought remains untested.

>>>>

A second fundamental question regarding forest responses to drought is how species' traits -- alone and in interaction with tree size -- influence drought response. Xylem architecture plays a role, with diffuse porous species tending to be more drought sensitive than ring-porous species [@kannenberg_linking_2019] *(Elliot et al. 2015-doi: 10.1111/gcb.13045, Friedrichs et al. 2009-doi:10.1007/s00468-009-0315-2)*, but does not differentiate species beyond broad classes. To link species' level drought responses to fundamental physiological characteristics, it is important to understand how traits shape drought responses. Commonly measured traits including wood density ($WD$) and leaf mass per area ($LMA$) have been linked to drought responses in temperate deciduous forests [@abrams_adaptations_1990; @guerfel_impacts_2009]*(Hoffmann et al. 2011- doi: 10.1111/j.1365-2486.2011.02401.x)* and other forest biomes around the world (*Greenwood et al. 2007- doi: 10.1111/ele.12748*). However, the direction of response is not always consistent; wood density correlated negatively with drought tree drought performance in a broadleaf deciduous forest in the southeast US *(Hoffmann et al. 2011)* but positively across a global scale (*Greenwood et al. 2007- doi: 10.1111/ele.12748*). Leaf hydraulic traits with more direct linkage to plant hydraulic function, including leaf area shrinkage upon dessication ($PLA_{dry}$; [@scoffoni_leaf_2014]) and turgor loss point ($\pi_{tlp}$)--i.e., the water potential at which leaf wilting occurs [@bartlett_correlations_2016], are emerging as traits with potential to explain greater variation in plant distribution and function than the more commonly-measured traits such as $WD$ and $LMA$ [@medeiros_extensive_2019]. The ability of these hydraulic traits to tree performance under drought remains untested.
